# Correlation 



## Visulization of pair-wise correlation in R



## Correlation and p-values of all combinations of all rows of two matrices

### Prepare the data

```{r}
bac1 <- c(1,2,3,4,5)
bac2 <- c(2,3,4,5,1)
bac3 <- c(4,5,1,2,3)
bac4 <- c(5,1,2,3,4)
bac <- as.data.frame(cbind(bac1, bac2, bac3, bac4 ))
colnames(bac) <- c("station1", "station2", "station3", "station4")
rownames(bac) <- c("bac1", "bac2", "bac3", "bac4", "bac5")

fac1 <- c(1,2,3,4,5,6)
fac2 <- c(2,3,4,5,1,6)
fac3<- c(3,4,5,1,2,1)
fac4<- c(4,5,1,2,3, 6)
fac <- as.data.frame(cbind(fac1, fac2, fac3, fac4))
colnames(fac) <- c("station1", "station2", "station3", "station4")
rownames(fac) <- c("fac1", "fac2", "fac3", "fac4", "fac5", "fac6")

bac
fac
```




### Another way to do this is to use `corr.test`

You can just pass the full matrices to the `cor` function (or psych::corr.test)and it takes care of finding the correlation of the relevant columns.



```{r}
#install.packages("psych")
library(reshape2)
#cor.eff<-cor(t(fac), t(bac))
#coreff <- reshape2::melt(cor.eff)
pval <- psych::corr.test(t(fac), t(bac), adjust="fdr")
pval
cor_pval<- merge(melt(pval$r, value.name="cor"), melt(pval$p, value.name="p-value"), by=c("Var1", "Var2"))
head(cor_pval)

```


### Read data

```
data/cor_data_metablomics.txt         
data/cor_data_transcriptome.txt

```


### Calculate the correlation (Advance)


```{r}
tbac <- data.frame(t(bac))
tfac <- data.frame(t(fac))

f <- function (x, y) cor(x, y)
pval <- function(x, y) cor.test(x, y)$p.val

tab <- outer(tfac, tbac, Vectorize(f))
tabp<-outer(tfac, tbac, Vectorize(pval))

as.data.frame.table(tabp)
```

### Reference

Correlation/p values of all combinations of all rows of two matrices: https://stackoverflow.com/questions/41793219/correlation-p-values-of-all-combinations-of-all-rows-of-two-matrices



## Different ways to visulize correlation

### Display correlation using `chart.Correlation` in `PerformanceAnalytics`

```{r}
library(datasets)
data(iris)
library("PerformanceAnalytics")
chart.Correlation(iris[, 1:4], histogram=TRUE, pch=19)
```

```{r}
library(corrplot)
corrplot.mixed(cor(iris[, 1:4]), order="original", tl.col="black")
```

## Reference


Seven Easy Graphs to Visualize Correlation Matrices in R: http://jamesmarquezportfolio.com/correlation_matrices_in_r.html

Example 9.17: (much) better pairs plots: https://www.r-bloggers.com/example-9-17-much-better-pairs-plots/


Introduction to Correlation: https://rpubs.com/aaronsc32/linear-relationship-pearson-r-correlation


Spearman Rank Correlation: https://rstudio-pubs-static.s3.amazonaws.com/191093_4169c5282eb145a491a5b1924941a6ba.html

