# T test

## Calculate t-statistic step by step in R


```{r}
#load dataset
library(reshape2)
data(tips)

```

```{r}
#create tip percent variable
tips$percent=tips$tip/tips$total_bill
```

```{r}
#Split dataset for ease
splits<-split(tips, tips$sex)

```

```{r}
#Save data sets
women<-splits[[1]] 
men<-splits[[2]]
```

```{r}
#variance by group sample size
var_women<-var(women$percent)/length(women$percent)
var_men<-var(men$percent)/length(men$percent)
```

```{r}
#Sum
total_variance<-var_women+var_men
```

```{r}
#Squre Root
sqrt_variance<-sqrt(total_variance)

#Group means by pooled variances
(mean(women$percent)-mean(men$percent))/sqrt_variance
#T.test
t.test(percent~sex, data=tips)
```


### 

Your data set is probably not the best illustrative example in terms of normality assumption... but anyway, here is some quick R code to reproduce some of the calculation of t.test().

#### Equal variances

```{r}
head(tips)
t.test(percent ~ sex, data=tips, var.equal=TRUE)

x1 <- tips$percent[tips$sex == "Female"]
x2 <- tips$percent[tips$sex == "Male"]
n1 <- length(x1)
n2 <- length(x2)

var.pooled <- weighted.mean(x=c(var(x1), var(x2)), w=c(n1 - 1, n2 - 1))

t <- (mean(x1) - mean(x2)) / sqrt(var.pooled / n1 + var.pooled / n2)
t
df <- n1 + n2 - 2
df
```




#### Unequal variance

```{r}
t.test(percent ~ sex, data=tips, var.equal=FALSE)
# Welch Two Sample t-test
# 
# data:  percent by sex
# t = 1.1433, df = 206.759, p-value = 0.2542
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#   -0.006404119  0.024084498
# sample estimates:
#   mean in group Female   mean in group Male 
# 0.1664907            0.1576505 

x1 <- tips$percent[tips$sex == "Female"]
x2 <- tips$percent[tips$sex == "Male"]
n1 <- length(x1)
n2 <- length(x2)

t <- (mean(x1) - mean(x2)) / sqrt(var(x1) / n1 + var(x2) / n2)
t
# [1] 1.143277
df.num <- (var(x1) / n1 + var(x2) / n2)^2
df.denom <- var(x1)^2 / (n1^2 * (n1 - 1)) + var(x2)^2 / (n2^2 * (n2 - 1))
df <- df.num / df.denom
df
# [1] 206.7587
```


## Calculate the standard deviation in R

$variance = \frac{sum((x-mean(x))^2)}{(length(x)-1)}$


```{r}
a <- c(179,160,136,227)
sd(a)
sqrt(sum((a-mean(a))^2/(length(a)-1)))
```


## Reference


Calculate t-statistic step by step in R: https://stats.stackexchange.com/questions/141593/calculate-t-statistic-step-by-step-in-r
