# MRG

Sir Francis Galton (1822–1911) was an English statistician. He founded many concepts in statistics, such as correlation, quartile, percentile and regression, that are still being used today.



## Read the data

Consider the data collected by Francis Galton in the 1880s, stored in a modern format in the `galton.csv` file. In this file, `heights` is the variable containing the child’s heights, while the `father`’s and `mother`’s height is contained in the variables father and mother.

The family variable is a numerical code identifying children in the same family; the number of kids in this family is in `nkids`.



```{r}
## Data from  https://github.com/thomas-haslwanter/statsintro_python/blob/master/ISP/Code_Quantlets/08_TestsMeanValues/anovaOneway/galton.csv
tab<-read.csv("data/galton.csv")
head(tab)
```

Check the number of rows and columns:

```{r}
dim(tab)
```

Covert the column of `sex` into numberic values:

```{r}
tab$sex=as.numeric(tab$sex) -1 
head(tab)
```
Remove the columns of `nkids`: 

```{r}
tab<-tab[, -c(6)]
head(tab)
```








