# Working with missing values

## 

```{r}
## Some sample data
set.seed(0)
dat <- matrix(1:100, 10, 10)
dat[sample(1:100, 10)] <- NA
dat <- data.frame(dat)

```




## Summaryize the missing values


### Summarize the missing values using `mice` package

The `mice` package provides a nice function md.pattern() to get a better understanding of the pattern of missing data. 



```{r}
#install.packages("mice")
library(mice)
md.pattern(dat)
```

### Summarize the missing values using `VIM` packages

```{r}
#install.packages("VIM")
library(VIM)
aggr(dat, numbers = TRUE, prop = c(TRUE, FALSE))
aggr(t(dat), numbers = TRUE, prop = c(TRUE, FALSE))

```

```{r}
aggr_plot <- aggr(dat, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(dat), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```

```{r}
tdat = t(dat)
aggr_plot <- aggr(tdat, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(tdat), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```


## Delete columns/rows with more that x% missing 

```{r}
## Remove columns with more than 50% NA
dat_file = dat[, -which(colMeans(is.na(dat)) > 0.5)]

```

## Imputing the data

### Imputing the data with row-wise mean

```{r}
dat.imp = dat
k <- which(is.na(dat.imp), arr.ind=TRUE)
dat.imp[k] <- rowMeans(dat.imp, na.rm=TRUE)[k[,1]]
```


### Imputing the data with  row-wise mean using `mice`

```{r}
tempDat <- mice(as.matrix(dat), method = "mean")
summary(tempDat)
impDat<- complete(tempDat)
impDat
```


## Reference

Imputing Missing Data with R; MICE package: https://datascienceplus.com/imputing-missing-data-with-r-mice-package/

Delete columns/rows with more that x% missing : https://stackoverflow.com/questions/31848156/delete-columns-rows-with-more-that-x-missing


Tutorial on 5 Powerful R Packages used for imputing missing values
: https://www.analyticsvidhya.com/blog/2016/03/tutorial-powerful-packages-imputing-missing-values/
