# Random forest



## Practical experiments


### Random forest for prediction of iris 


The `caret` package (short for _C_lassification _A_nd _RE_gression _T_raining) is a set of functions that attempt to streamline the process for creating predictive models. The package contains tools for:

* data splitting
* pre-processing
* feature selection
* model tuning using resampling
* variable importance estimation

#### Required R packages

Required packages:

```
caret   
AppliedPredictiveModeling 
ellipse 
```

Attribute Information:

1. sepal length in cm 

2. sepal width in cm 

3. petal length in cm 

4. petal width in cm 

5. class: 

-- Iris Setosa 
-- Iris Versicolour 
-- Iris Virginica

```{r}
str(iris)
head(iris)
dim(iris)
```

#### Visulization

##### Scatterplot Matrix

```{r}
#install.packages("AppliedPredictiveModeling")
library(AppliedPredictiveModeling)
transparentTheme(trans = .4)
#install.packages("caret")
library(caret)
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "pairs",
            ## Add a key at the top
            auto.key = list(columns = 3))
```

##### Scatterplot Matrix with Ellipses

```{r}
#install.packages("ellipse")
library(ellipse)
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "ellipse",
            ## Add a key at the top
            auto.key = list(columns = 3))
```

##### Overlayed Density Plots

```{r}
featurePlot(x = iris[, 1:4], 
            y = iris$Species,
            plot = "density", 
            ## Pass in options to xyplot() to 
            ## make it prettier
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")), 
            adjust = 1.5, 
            pch = "|", 
            layout = c(4, 1), 
            auto.key = list(columns = 3))
```

##### Box Plots

```{r}
featurePlot(x = iris[, 1:4], 
            y = iris$Species, 
            plot = "box", 
            ## Pass in options to bwplot() 
            scales = list(y = list(relation="free"),
                          x = list(rot = 90)),  
            layout = c(4,1 ), 
            auto.key = list(columns = 2))
```
 
 
##### Machine leraning - Random forest

##### Model Train 

```{r}
set.seed(186)
# Data Splitting
train_index <- createDataPartition(iris$Species, p = 0.75, , times=1, list = FALSE)
train_set = iris[train_index, ]
test_set  = iris[-train_index, ]
fit_rf_cv <- train(Species ~ ., data=train_set, method='rf', metric = "Accuracy",
                   trControl=trainControl(method="cv",number=5)) 
fit_rf_cv
## Variance importance
rfVarImpcv = varImp(fit_rf_cv)
rfVarImpcv
```


##### Testing test data

```{r}
## predict
test_set$predict_rf <- predict(fit_rf_cv, test_set, "raw")
confusionMatrix(test_set$predict_rf, test_set$Species)
```


