# Partial least square analysis

## Prepare the data


```{r}
library(ropls)
data(sacurine)
names(sacurine)
```


We attach sacurine to the search path and display a summary of the content of the dataMatrix, sampleMetadata and variableMetadata with the strF Function of the ropls package (see also str):

```{r}
attach(sacurine)
strF(dataMatrix)
head(dataMatrix)
```


```{r}
sampleMetadata
```


```{r}
summary(sampleMetadata)
```

## Principal Component Analysis (PCA)

We perform a PCA on the dataMatrix matrix (samples as rows, variables as columns), with the opls method:

```{r}
sacurine.pca <- opls(dataMatrix)

```


```{r}
genderFc <- sampleMetadata[, "gender"]
plot(sacurine.pca, typeVc = "x-score",
parAsColFcVn = genderFc, parEllipsesL = TRUE)

```

## Partial least-squares: PLS and PLS-DA

For PLS (and OPLS), the Y response(s) must be provided to the opls method. Y can be either a numeric vector (respectively matrix) for single (respectively multiple) (O)PLS regression, or a character factor for (O)PLS-DA classification as in the following example with the gender qualitative response:

```{r}
sacurine.plsda <- opls(dataMatrix, genderFc)

```


```{r}
genderFc
```

```{r}
head(dataMatrix)
```
## Reference 

https://bioconductor.org/packages/release/bioc/vignettes/ropls/inst/doc/ropls-vignette.html#4_hands-on
