---
title: 'Principal Component Analysis'
author: 'Dataset new_data'
output:
  pdf_document: default
---
This dataset contains 150 individuals and 2 variables.

- - -

```{r, echo = FALSE}
load('Workspace.RData')
```
### 1. Study of the outliers
The analysis of the graphs does not detect any outlier.

- - -

### 2. Inertia distribution
The inertia of the first dimensions shows if there are strong relationships between variables and suggests the number of dimensions that should be studied.

The first two dimensions of PCA express **100%** of the total dataset inertia ; that means that 100% of the individuals (or variables) cloud total variability is explained by the plane.
The inertia observed on the first plane is smaller than the reference value that equals **100%**, therefore low in comparison
(the reference value is the 0.95-quantile of the inertia percentages distribution obtained by simulating 1429 data tables of equivalent size on the basis of a normal distribution).
However, the inertia related to the first dimension is greater than the reference value **57.75%**.
Even if the inertia projected on the first plane is not significant, these explained by the first dimension is significant.

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
par(mar = c(2.6, 4.1, 1.1, 2.1))
barplot(res$eig[,2], names.arg = 1:nrow(res$eig))
```

**Figure 2 - Decomposition of the total inertia on the components of the PCA**
*The first factor is largely dominant: it expresses itself 98.14% of the data variability.*
*Note that in such a case, the variability related to the other components might be meaningless, despite of a high percentage.*

An estimation of the right number of axis to interpret suggests to restrict the analysis to the description of the first 1 axis.
These axis present an amount of inertia greater than those obtained by the 0.95-quantile of random distributions (98.14% against 57.75%).
This observation suggests that only this axis is carrying a real information.
As a consequence, the description will stand to these axis.

- - -

### 3. Description of the dimension 1

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("115", "135", "142", "108", "123", "146", "130", "145", "116", 
"137")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = 1:2, choix = 'ind', invisible = 'quali', title = '', cex = cex)
```

**Figure 3.1 - Individuals factor map (PCA)**
*The labeled individuals are those with the higher contribution to the plane construction.*

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("petal_length", "petal_width")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.PCA(res, select = drawn, axes = 1:2, choix = 'var', title = '', cex = cex)
```

**Figure 3.2 - Variables factor map (PCA)**
*The labeled variables are those the best shown on the plane.*

* * *

The **dimension 1** opposes individuals such as *108*, *115*, *116*, *123*, *130*, *135* and *142* (to the right of the graph, characterized by a strongly positive coordinate on the axis)
to individuals characterized by a strongly negative coordinate on the axis (to the left of the graph).

The group in which the individuals *108*, *115*, *116*, *123*, *130*, *135* and *142* stand (characterized by a positive coordinate on the axis) is sharing :

- high values for the variables *petal_width* and *petal_length* (variables are sorted from the strongest).

The group 2 (characterized by a negative coordinate on the axis) is sharing :

- low values for the variables *petal_length* and *petal_width* (variables are sorted from the weakest).

Note that the variables *petal_length* and *petal_width* are highly correlated with this dimension (respective correlation of 0.98, 0.98). These variables could therefore summarize themselve the dimension 1.

- - -

### 4. Classification

```{r, echo = FALSE}
res.hcpc = HCPC(res, nb.clust = -1, graph = FALSE)
```

```{r, echo = FALSE, fig.align = 'center', fig.height = 3.5, fig.width = 5.5}
drawn <-
c("115", "135", "142", "108", "123", "146", "130", "145", "116", 
"137")
par(mar = c(4.1, 4.1, 1.1, 2.1))
plot.HCPC(res.hcpc, choice = 'map', draw.tree = FALSE, select = drawn, title = '')
```

**Figure 4 - Ascending Hierarchical Classification of the individuals.**
*The classification made on individuals reveals 3 clusters.*


The **cluster 1** is made of individuals sharing :

- low values for the variables *petal_length* and *petal_width* (variables are sorted from the weakest).

The **cluster 2** is made of individuals such as *135*. This group is characterized by :

- high values for the variable *petal_length*.

The **cluster 3** is made of individuals such as *108*, *115*, *116*, *123*, *130*, *137*, *142*, *145* and *146*. This group is characterized by :

- high values for the variables *petal_width* and *petal_length* (variables are sorted from the strongest).

- - -

## Annexes
```{r, comment = ''}
dimdesc(res, axes = 1:1)
```
**Figure 5 - List of variables characterizing the dimensions of the analysis.**



```{r, comment = ''}
res.hcpc$desc.var
```
**Figure 6 - List of variables characterizing the clusters of the classification.**


