# Unstanding power in statistical test


## Calculating the power using a normal distribution


We will assume that the standard deviation is 2, and the sample size is 20. In the example below we will use a 95% confidence level and wish to find the power to detect a true mean that differs from 5 by an amount of 1.5. (All of these numbers are made up solely for this example.) The commands to find the confidence interval in R are the following:


```{r}
a <- 5 # 
sd <- 2 # standard deviation
n <- 20 # sample size
error <- qnorm(0.975)*sd/sqrt(n)
left <- a-error
right <- a+error
error 
left
right
```


Next we find the Z-scores for the left and right values assuming that the true mean is 5+1.5=6.5:

```{r}
assumed = a + 1.5
Zleft <- (left-assumed)/(sd/sqrt(n))
Zright <-(right-assumed)/(sd/sqrt(n))
p <- pnorm(Zright)-pnorm(Zleft)
p
1-p
```

```{r}
Zleft
Zright
```

In this example, the power of the test is approximately 91.8%. If the true mean differs from 5 by 1.5 then the probability that we will reject the null hypothesis is approximately 91.8%.

```{r}
se = sqrt(sd/n)
alpha = 0.05
cuts = c(1-alpha/2, alpha/2)
crits = qnorm(cuts, 5, se)    ## Critical Values Based on Null

lowlim = a - 4*sd
uplim  = a + 4*sd

### Distribution normal mean 5
x.grid = seq(lowlim,uplim, length= 500)
dens.all = dnorm(x.grid, mean=a, sd = sd)
plot(x.grid, dens.all, type="l", xlab="X", ylab="Density")

pcts = c(0.025,0.975)
outside=qnorm(pcts,a,sd)
below = min(outside)
above = max(outside)

x.below    = x.grid[x.grid<below]
dens.below = dens.all[x.grid<below]
polygon(c(x.below,rev(x.below)),c(rep(0,length(x.below)),rev(dens.below)),col="grey",density=40)
x.above    = x.grid[x.grid>above]
dens.above = dens.all[x.grid>above]
polygon(c(x.above,rev(x.above)),c(rep(0,length(x.above)),rev(dens.above)),col="grey",density=40)


### 

mu1 = a + 1.5
lowlim1 = mu1 - 4*sd
uplim1 =  mu1 + 4*sd
x.grid = seq(lowlim1, uplim1, length= 500)
dens.all = dnorm(x.grid, mean=mu1, sd = sd)
lines(x.grid,dens.all, col="blue", lty="dashed")

outside=qnorm(pcts,mu1,sd)
below = min(outside)
above = max(outside)

x.below    = x.grid[x.grid<below]
dens.below = dens.all[x.grid<below]
polygon(c(x.below,rev(x.below)),c(rep(0,length(x.below)),rev(dens.below)),col="blue",density=40)
x.above    = x.grid[x.grid>above]
dens.above = dens.all[x.grid>above]
polygon(c(x.above,rev(x.above)),c(rep(0,length(x.above)),rev(dens.above)),col="blue",density=40)

## 
abline(v = left,  lty="dashed", col="red")
abline(v = right, lty="dashed", col="red")

```

```{r}

```

## Function `shadenorm()`

```{r}
## Reference: https://gist.github.com/tony91782/950812
## -------------------------- ##
## An R Function that Shades  ##
## under a normal density.    ##
##                            ##
## This is a convenience      ##
## function for polygon()     ##
## -------------------------- ##

shadenorm = function(below=NULL, above=NULL, pcts = c(0.025,0.975), mu=0, sig=1, numpts = 500, color = "gray", dens = 40,
                   lines=FALSE,between=NULL,outside=NULL){
   if(is.null(between)){
      bothnull = is.null(below) & is.null(above)
      if(bothnull==TRUE){
        below = ifelse(is.null(below), qnorm(pcts[1],mu,sig), below)
        above = ifelse(is.null(above), qnorm(pcts[2],mu,sig), above)
      }
   }

   if(is.null(outside)==FALSE){
        if(is.numeric(outside)==FALSE){if(outside==TRUE){outside=qnorm(pcts,mu,sig)}}
        below = min(outside)
        above = max(outside)
   }

   lowlim = mu - 4*sig
   uplim  = mu + 4*sig

   x.grid = seq(lowlim,uplim, length= numpts)
   dens.all = dnorm(x.grid,mean=mu, sd = sig)

   if(lines==FALSE){
         plot(x.grid, dens.all, type="l", xlab="X", ylab="Density")
   }

   if(lines==TRUE){
         lines(x.grid,dens.all)
   }

   if(is.null(below)==FALSE){
       x.below    = x.grid[x.grid<below]
       dens.below = dens.all[x.grid<below]
       polygon(c(x.below,rev(x.below)),c(rep(0,length(x.below)),rev(dens.below)),col=color,density=dens)
   }

   if(is.null(above)==FALSE){
       x.above    = x.grid[x.grid>above]
       dens.above = dens.all[x.grid>above]
       polygon(c(x.above,rev(x.above)),c(rep(0,length(x.above)),rev(dens.above)),col=color,density=dens)
   }

   

   if(is.null(between)==FALSE){
        if(is.numeric(between)==FALSE){if(between==TRUE){between=qnorm(pcts,mu,sig)}}
        from = min(between)
        to   = max(between)

        x.between    = x.grid[x.grid>from&x.grid<to]
        dens.between = dens.all[x.grid>from&x.grid<to]
        polygon(c(x.between,rev(x.between)),c(rep(0,length(x.between)),rev(dens.between)),col=color,density=dens)
   }
}
```



```{r}
## Parameters of the Problem ##

mu0   = 6     ## Null-hypothesized value
alpha = 0.05  ## Significance Level
s.sq  = 16    ## Variance from pilot study
n1    = 64    ## Sample Size
alt =  6.5

se = sqrt(s.sq/n1)

## -------------------------- ##
## What is Statistical Power? ##
## -------------------------- ##

cuts = c(1-alpha/2, alpha/2)
crits = qnorm(cuts, mu0, sqrt(s.sq/n1))    ## Critical Values Based on Null
shadenorm(mu=mu0, sig = se, outside=crits) ## My own function
shadenorm(mu = 6.5, sig = se, lines=TRUE, outside=crits, col="blue")
```

## Write a Function to Compute it

```{r}
power = function(theta, mu, var, n, alpha=0.05){
   ## The qnorm function is simply the inverse of the cdf
   ## You can think of as the inverse of pnorm
   crit.l = qnorm(alpha/2, mu, sqrt(var/ n))    ## Critical Value Based on Null
   crit.h = qnorm(1-alpha/2, mu, sqrt(var/ n))  ## Critical Value Based on Null
   
   # pnorm is the R function that calculates the c.d.f. (Cumulative distribution function)
   # P(X < ctrit.l)
   pr.high = pnorm(crit.h, theta, sd = sqrt(var/ n),lower.tail=FALSE) ## Prob Reject High
   ## P(X > ctrit.l)
   pr.low  = pnorm(crit.l, theta, sd = sqrt(var/ n))                  ## Prob Reject Low
   pow = pr.low+pr.high

   pow
}
```

```{r}
power(thet=6.5, mu=6, var= 16, n = 64)
thet = seq(3,9, by=0.01)
pow = power(thet, mu0, s.sq, n1)
plot(thet, pow, type="l", ylim=c(0,1), main="My Power Plot")

n2 = 256  ## New Larger Sample Size
pow2 = power(thet, mu0, s.sq,n2)
lines(thet, pow2, lty="dashed", col="blue")
abline(h=0.05, lty="dotted")
```


# Reference

An R Function that Shades under a normal density: https://gist.github.com/tony91782/950812

Computing statistical power: https://www.r-bloggers.com/computing-statistical-power/

Probability Distributions in R: http://www.stat.umn.edu/geyer/old/5101/rlook.html

Introduction to dnorm, pnorm, qnorm, and rnorm for new biostatisticians: http://seankross.com/notes/dpqr/

Calculating The Power Of A Test: https://www.cyclismo.org/tutorial/R/power.html
